Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-11-18 demystify.tex
updates

In recent years many universities start requiring a statement about diversity, equity, and inclusion (DEI) as part of the application (also see \autoref{sec:urm}). This has caused confusion among many applicants, especially international students. They often wonder what to write about, especially if they are not from a minority group, have not experienced anything related the topic, or even not sure what it is about.

A DEI statement is about showing your \emph{awareness of DEI}, your understanding of the challenges others face, and actions you've taken to contribute. Customize what you write to your own background and experiences.  Here are some tips:

@@

\item \textbf{Show your awareness, customized to your own experiences}. For example, if you come from a different culture or country, talk about DEI here, e.g., many countries completely lack diversity in CS programs, talk about it, what does minority group face in your country?  How do you think DEI can benefit?

@@

Moreover, do not make up stories or exaggerate your experiences.  Just show your awareness and understanding of the topic and what you did or how you plan to contribute to it (e.g., mentoring students, helping others).  One of the best DEI essay I have read start the acknowledgement that the applicant is not a minority, but a white male in a STEM field (i.e., that make them the majority), and that they do not have much experience in DEI, but show that they are aware of the challenges faced by minority groups and how they can contribute to it.
\end{commentbox}

>>>>

In recent years many universities start requiring a statement about diversity, equity, and inclusion (DEI) as part of the application (also see \autoref{sec:urm}). Students, especially international ones, often wonder what to write about, especially if they are not from a minority group, have not experienced anything related the topic, or even not sure what it is about.

A DEI statement is about showing your \emph{awareness} of DEI, \emph{understanding} of the challenges others face, and \emph{potential things} you can do to contribute.

@@

\item \textbf{Show your awareness, customized to your own experiences}. If you come from a different culture or country, talk about DEI there. Many countries, especially developing countries, do not have/care about DEI or even have laws against it, so does minority group face in your country?  How do you think DEI can benefit? Remember, you would have a unique perspective on this topic that we might not know about, so use that to your advantage.

@@

Moreover, do not make up stories or exaggerate your experiences.  Just show your awareness and understanding of the topic and what you did or how you plan to contribute to it (e.g., mentoring students, helping others). One of the best DEI statements I have read starts the acknowledgement that the applicant is not a minority, but a white male in a STEM field (i.e., that make them the majority), and that they do not have much experience in DEI, but show that they are aware of the challenges faced by minority groups and how they can contribute to it.
\end{commentbox}
